Testing Charts: viewers' perceptual accuracy in surveys

### 2023-07-31 SDSS-2023/testing-charts-accuracy.qmd
adding equation numbers

When combining responses from different surveys, weights are rescaled with respect to the total population, so that their order after combining still reflects their relative importance. We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ... $S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
$\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
$$
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
$$ 
where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
Here, $d_i$ are estimated as 
$$
d_i = 1 + CV(w \in S_i)^2
$$ 
where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 : 
$$ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
$$

@@

$$
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
$$

@@

$$
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
$$

@@

Using linear scores for the response of certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty by using a cell-means model of the form: 
$$
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
$$ 
where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
We also assume that errors are normally distributed, i.e. $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
The results are shown in @fig-certainty.
We find that the highest scores for certainty are associated with the aligned task, but certainty scores are not significantly different between correct and wrong responses.
Scores are (mostly) significantly lower for the unaligned task.
Interestingly, the lowest certainty scores are associated with correct responses to the unaligned task; although respondents successfully complete the task, they are less certain in their selection.

@@

$$
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
$$

>>>>

When combining responses from different surveys, weights are rescaled with respect to the total population, so that their order after combining still reflects their relative importance. We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ... $S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
$\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
\begin{equation}
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\end{equation}
where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
Here, $d_i$ are estimated as 
\begin{equation}
d_i = 1 + CV(w \in S_i)^2
\end{equation}
where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 : 
\begin{equation}
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\end{equation}

@@

\begin{equation}
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
\end{equation}

@@

\begin{equation}
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
\end{equation}

@@

Using linear scores for the response of certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty by using a cell-means model of the form: 
\begin{equation}
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
\end{equation}
where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
We also assume that errors are normally distributed, i.e. $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
The results are shown in @fig-certainty.
We find that the highest scores for certainty are associated with the aligned task, but certainty scores are not significantly different between correct and wrong responses.
Scores are (mostly) significantly lower for the unaligned task.
Interestingly, the lowest certainty scores are associated with correct responses to the unaligned task; although respondents successfully complete the task, they are less certain in their selection.

@@

\begin{equation}
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
\end{equation}

### 2023-07-31 SDSS-2023/testing-charts-accuracy.tex
adding equation numbers

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication.
Decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

When combining responses from different surveys, weights are rescaled
with respect to the total population, so that their order after
combining still reflects their relative importance. We \emph{combine}
(rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
\ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as described in
O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
\(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
\(\sum_i \lambda_i = 1\) is given as

\[
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\]

where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
the design effects for the estimators. Here, \(d_i\) are estimated as

\[
d_i = 1 + CV(w \in S_i)^2
\]

where \(CV\) is the coefficient of variation of the weights \(w\) within
each sample, and is estimated as in Kish (1965) :

\[ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\]

@@

\[
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
\]

@@

\[
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
\]

The resulting model estimates and confidence intervals are displayed in
\autoref{tbl-zoom-123}. We observe that zooming behavior \emph{only}
differs significantly by structural design; rates do not differ
significantly between the aligned and unaligned tasks, nor do they
differ significantly by the chosen response. This higher rate of zooming
does not necessarily lead to higher accuracy; although rates of accuracy
are significantly higher for the vertical orientation relative to the
horizontal orientation, they are not significantly higher -- and in
fact, are significantly lower on the aligned task -- than the accuracy
for the horizontal wide orientation.
\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}

Using linear scores for the response of certainty, with `not certain at
all' assigned a score of 1 and `extremely certain' assigned a score of
5, we can estimate the effects of task, chart design, and correctness on
certainty by using a cell-means model of the form:

\[
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
\]

where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
(measured on a scale from 1 to 5) of the four combinations of task and
correctness by each of the three designs, where \(i = 1, 2\) encodes
unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
\(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
stacked bar charts, respectively. We also assume that errors are
normally distributed,
i.e.~\(\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)\) for all
\(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
We find that the highest scores for certainty are associated with the
aligned task, but certainty scores are not significantly different
between correct and wrong responses. Scores are (mostly) significantly
lower for the unaligned task. Interestingly, the lowest certainty scores
are associated with correct responses to the unaligned task; although
respondents successfully complete the task, they are less certain in
their selection.

@@

\[
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
\]

>>>>

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication.
Decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

When combining responses from different surveys, weights are rescaled
with respect to the total population, so that their order after
combining still reflects their relative importance. We \emph{combine}
(rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
\ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as described in
O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
\(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
\(\sum_i \lambda_i = 1\) is given as \begin{equation}
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\end{equation} where \(n_i\) is the nominal sample of survey \(S_i\) and
\(d_i\) are the design effects for the estimators. Here, \(d_i\) are
estimated as \begin{equation}
d_i = 1 + CV(w \in S_i)^2
\end{equation} where \(CV\) is the coefficient of variation of the
weights \(w\) within each sample, and is estimated as in Kish (1965) :
\begin{equation}
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\end{equation}

@@

\begin{equation}
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
\end{equation}

@@

\begin{equation}
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
\end{equation}

The resulting model estimates and confidence intervals are displayed in
\autoref{tbl-zoom-123}. We observe that zooming behavior \emph{only}
differs significantly by structural design; rates do not differ
significantly between the aligned and unaligned tasks, nor do they
differ significantly by the chosen response. This higher rate of zooming
does not necessarily lead to higher accuracy; although rates of accuracy
are significantly higher for the vertical orientation relative to the
horizontal orientation, they are not significantly higher -- and in
fact, are significantly lower on the aligned task -- than the accuracy
for the horizontal wide orientation.

Using linear scores for the response of certainty, with `not certain at
all' assigned a score of 1 and `extremely certain' assigned a score of
5, we can estimate the effects of task, chart design, and correctness on
certainty by using a cell-means model of the form: \begin{equation}
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
\end{equation} where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average
certainty (measured on a scale from 1 to 5) of the four combinations of
task and correctness by each of the three designs, where \(i = 1, 2\)
encodes unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
\(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
stacked bar charts, respectively. We also assume that errors are
normally distributed,
i.e.~\(\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)\) for all
\(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
We find that the highest scores for certainty are associated with the
aligned task, but certainty scores are not significantly different
between correct and wrong responses. Scores are (mostly) significantly
lower for the unaligned task. Interestingly, the lowest certainty scores
are associated with correct responses to the unaligned task; although
respondents successfully complete the task, they are less certain in
their selection.

@@

\begin{equation}
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
\end{equation}
